# Dimensionality reduction exercise
## The data
### 1. Load and describe data
Code for data wrangling can be found [here](https://github.com/Juhous/IODS-project/blob/master/create_human.R)

```{r, message = FALSE}
#Hidden
source("helper_functions.R")
library(magrittr)
library(GGally)
library(stringr)
library(knitr)
library(kableExtra)
library(ggplot2)
library(tidyr)
library(MASS)
library(dplyr)
```
```{r}
df <- read.csv(file = "data/human.csv", row.names = 1) 
glimpse(df)
```
Dataset contains information of human development charasteritics. It totals 189 observations (countries) and 8 variables. 

### 2. Graphical overview of data
```{r, out.width = "100%"}
#Hidden
df %>% summaryKable() %>% 
  kable("html", align = "rrr", caption = "Data variable summary") %>%
  kable_styling(bootstrap_options = c("hover", "condensed")) %>% 
  scroll_box(width = "100%")
```
```{R, warning = F, out.width = "100%"}
# Hidden
ggpairs(df,
  title = "Study variable overview",
  upper = list(continuous = wrap("cor", size = 3)),
  lower = list(
    continuous = wrap("points", alpha = .2, size = .4),
    combo = wrap("facethist", bins = 20))) +
theme(axis.text.x = element_text(
                  angle = 90,
                  color = "black",
                  size = 6,
                  vjust = .5),
      axis.text.y = element_text(color = "black", size = 6))
```
Many of the variables are normally distributed, but with eduRatio, GNIC and matMort, some problems might arise if normality is assumed. 
```{r}
#Hidden
cor(df, use = "pairwise.complete.obs") %>% 
  corrplot::corrplot(method = "number", type = "upper", 
                     tl.pos = "td", tl.srt = 25, 
                     tl.offset = 1, tl.cex = .7,
                     number.cex = .75, 
                     title = "Correlation between study variables",
                     mar=c(0,0,1,0),
                     diag = F) 
```

According to above plot, education correlates strongly and positively with life expectancy, and negatively with maternal mortality and adolescent births. Maternal mortality correlates strongly and negatively with education ratio, life expectancy and education. 

## Principal component analysis
### 3. PCA on non-normalized data
```{r}
pca <- prcomp(df)
pca %>% summary()
```
```{r, warning=F}
#Hidden
biplot(pca, choices = 1:2)
```

### 4. PCA on normalized data
```{r}
df_scaled <- scale(df)
pca <- prcomp(df_scaled)
pca %>% summary
```
```{r, warning=F}
#Hidden
biplot(pca, choices = 1:2)
```

### 5. Interpret differences

## 6. MCA with tea dataset
```{r}
library(FactoMineR)
data(tea)
df <- tea

mca <- df %>% select(breakfast, friends, resto, Tea, Sport) %>%
  MCA(graph = FALSE)
mca %>% summary
```
```{r}
#Hidden
mca %>% plot(invisible=c("ind"))
```


